Paper

### 2017-11-13 Paper.Rmd
...

```{r echo=F, warning=F, message=F, eval=T, fig.width=7, fig.height=5}
chr.lengths = get.chr.lengths()
chr.lengths$chrom = sub('chr','',chr.lengths$chrom)
chr.lengths = subset(chr.lengths, chrom %in% c(1:22))

@@

globalCNdf = locs
globalCNdf = cbind(globalCNdf, t(dysplasia.df[,grep('^\\d+:', colnames(dysplasia.df), value=T)]))
globalCNdf = merge(globalCNdf, chr.lengths[,c('chrom','chr.length')], by.x='chr', by.y='chrom')
globalCNdf = melt(globalCNdf, id.vars=c('chr','start','end', 'chr.length'))
globalCNdf$Status = ifelse(globalCNdf$variable %in% names(which(labels == 1)), 'Progressor', 'Non-Progressor')

globalCNdf = merge(globalCNdf, patient.info[,c('Hospital.Research.ID','Patient','Samplename','Pathology')], by.x='variable', by.y='Samplename')

#globalCNdf = subset(globalCNdf, value != 0)

cn = function(x) {
  str = 'neutral'
  if(x <= -1) {
    str = 'loss'
  } else if (x >= 1) {
    str = 'gain'
  }
  return(str)
}

globalCNdf$cn = factor(sapply(globalCNdf$value, cn), levels=c('loss','neutral','gain'), ordered = T)
globalCNdf$cn = factor(sapply(globalCNdf$value, cn), levels=c('loss','neutral','gain'), ordered = T)
globalCNdf$cnB = factor(as.integer(globalCNdf$cn != 'neutral'))

cn.colors = c('orange3','grey88','steelblue2')

@@

# Average cn number...looks the same as expected
globalCNdf = globalCNdf %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
globalCN.plot = 
  ggplot(globalCNdf, aes(x=chr.length)) + facet_grid(Status~chr, scales='free_x', space='free_x', labeller=labeller(Status=n.status, chr=label_value)) +
  geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=cn), alpha=0.5, show.legend = T) + 
  scale_color_manual(values=cn.colors) + 
  scale_fill_manual(values=cn.colors) +
  labs(x='', y='Mean adjusted segmentation value', title='All pathologies') +
  theme(text=element_text(face='bold', size=14), axis.text.x=element_blank(), legend.position='bottom', 
        panel.grid = element_blank(), panel.background = element_blank(),strip.background =element_rect(fill="lightcyan2"),  
        panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines') )

cnLegend = get.legend(globalCN.plot)
globalCN.plot = globalCN.plot + theme(legend.position = 'none')

@@

globalCNBEdf = subset(globalCNdf, Pathology == 'BE') %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
globalBECN.plot = 
  ggplot(globalCNBEdf, aes(x=chr.length)) + 
  facet_grid(Status~chr, scales='free_x', space='free_x', labeller=labeller(Status=nb.status, chr=label_value)) + 
  #ylim(range(globalCNdf$value)) +
  geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=cn), alpha=0.5, show.legend = T) + 
  scale_color_manual(values=cn.colors) + scale_fill_manual(values=cn.colors) +
  labs(x='', y='Mean adjusted segmentation value', title='NDBE samples only') +
  theme(text=element_text(face='bold', size=14), axis.text.x=element_blank(), legend.position='bottom', 
        panel.grid = element_blank(), panel.background = element_blank(),strip.background =element_rect(fill="lightcyan2"),  
        panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines') )

@@

dfpts = subset(globalCNdf, Patient %in% c('57','76'))
dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )

pt.status <- c('57'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
              '76'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))

genomePtCN.plot = 
ggplot(dfpts, aes(x=chr.length)) + 
  facet_grid(Patient~chr, scales='free_x',space='free_x',  labeller = labeller(Patient = pt.status )) + 
  geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=cn), show.legend = F, alpha=0.5) + 
  scale_fill_manual(values=cn.colors) +
  labs(x='', y='Mean adjusted segmentation value') +
  theme(text=element_text(face='bold', size=14), axis.text.x=element_blank(), legend.position='bottom', 
        panel.grid = element_blank(), panel.background = element_blank(),strip.background =element_rect(fill="lightcyan2"),  
        panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines') )

dfpts = subset(globalCNdf, Patient %in% c('57','76') & Pathology == 'BE')
dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )

ptbe.status <- c('57'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
              '76'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
genomePtBECN.plot = 
  ggplot(dfpts, aes(x=chr.length)) + 
  facet_grid(Patient~chr, scales='free_x', space='free_x', labeller = labeller(Patient = ptbe.status )) + 
  geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=cn), show.legend = F, alpha=0.5) + 
  scale_fill_manual(values=cn.colors) +
  labs(x='', y='Mean adjusted segmentation value') +
  theme(text=element_text(face='bold', size=14), axis.text.x=element_blank(), legend.position='bottom', 
        panel.grid = element_blank(), panel.background = element_blank(),strip.background =element_rect(fill="lightcyan2"),  
        panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines') )

genomePtCN.plot 
genomePtBECN.plot
```

@@

```{r, echo=F, fig.height=5, fig.width=5, eval=F}
mp = model.performance(performance.at.1se, coefs)
mp$plot + labs(title=paste('No HGD/IMC samples,', folds, 'folds,', splits, 'patient splits'))

@@

```{r, echo=F, fig.height=5, fig.width=5, eval=F}
mp = model.performance(performance.at.1se, coefs)
mp$plot + labs(title=paste('No LGD/HGD/IMC samples,', folds, 'folds,', splits, 'patient splits'))

@@

ggplot(classification.rate, aes(model, mean, group=model, color=model)) + 
  geom_text( aes(label=round(mean, 3)), hjust=-0.5, size=5, color='gray39', show.legend = F) +
  geom_text( aes(y=mean+sme*3, label=paste(all.coef, "\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme),color='grey39', width=0.1, size=1) + 
  geom_point(size=4) +
  labs(x='', y='Classificaiton rate', title=paste('Classification rate per model at', select.alpha)) + 
  plot.theme + theme(axis.text.x=element_text(angle=45,hjust=1,face="bold",size=12), legend.position = 'none')

@@

custom.max<-function(cf, hazards, ymax) {
  print(cf)
  if (cf > 0) {
    max = hazards*ymax
    max = ifelse(max > ymax, ymax, max)
  } else {
    max = hazards*ymax
    max = ifelse(max < ymax, ymax, max)
  }
  return(max)
}
chh = chh %>% rowwise() %>% mutate('max'=custom.max(coef, hazards, ymax))

globalCN.plot + geom_rect(data=chh, aes(xmin=start, xmax=end, ymin=0, ymax=max, color=gl), alpha=0, show.legend=F) + scale_color_manual(values=c('goldenrod2','blue3'))
globalBECN.plot + geom_rect(data=chh, aes(xmin=start, xmax=end, ymin=0, ymax=max, color=gl), alpha=0, show.legend=F) + scale_color_manual(values=c('goldenrod2','blue3'))

ch$ymax = ifelse(ch$coef > 0, max(genomePtCN.plot$data$value), min(genomePtCN.plot$data$value))
chh = chh %>% rowwise() %>% mutate('max'=custom.max(coef, hazards, ymax))
genomePtCN.plot + geom_rect(data=chh, aes(xmin=start, xmax=end, ymin=0, ymax=max, color=gl), alpha=0, show.legend=F) + scale_color_manual(values=c('goldenrod2','blue3'))
genomePtBECN.plot + geom_rect(data=chh, aes(xmin=start, xmax=end, ymin=0, ymax=max, color=gl), alpha=0, show.legend=F) + scale_color_manual(values=c('goldenrod2','blue3'))

@@

p3 = ggplot(chh, aes(x=label,y=coef,fill=gl)) + geom_bar(stat='identity') + labs(x='Effects', y='') +
  coord_flip() + scale_fill_manual(values=c('darkgreen','darkred')) +
  plot.theme + theme(axis.line.x=element_blank(), axis.ticks=element_blank(),axis.text=element_blank(), panel.grid=element_blank(),panel.grid.major=element_blank(), legend.position='none')
p3

#  lay <- rbind(c(1,1,1,4),
#               c(2,2,2,5),
#               c(3,3,3,6),
#               c(7,7,7,8))
# 
# grid.arrange(p3,p1a,p1b,gp,p2a,p2b,p1leg,gp, layout_matrix=lay, widths=4:1, heights=c(2,4,4,1))
```

@@

```{r leaveoneout, echo=F, message=F, warning=F}
file = paste(cache.dir, 'loo.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop(paste("Missing model file", file))
}

df = as.data.frame(performance.at.1se)
ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
  geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
  geom_label(data=subset(df, performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
  labs(y='performance at lambda.1se', x='', title=paste('Performance for nested leave one out, at alpha=',select.alpha,sep='')) +
  plot.theme
```

@@

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=8, fig.width=4}
roc.plot <- function(roc,title="") {
  df = cbind.data.frame('Specificity'=rev(roc$specificities), 'Sensitivity'=rev(roc$sensitivities))

@@

```{r p53roc, echo=F, fig.height=5, fig.width=5}
p53samples = subset(patient.info, !is.na(p53.Status), select=c('Samplename','p53.Status','Status'))
p53roc = pROC::roc(as.numeric(p53samples$Status)-1, as.integer(p53samples$p53.Status)-1)

@@

```{r pathroc, echo=F, fig.height=5, fig.width=5 }
#sum(table(subset(patient.info, Status == 'P')$Pathology)[c('IMC','HGD')])/sum(table(subset(patient.info, Status == 'P')$Pathology))

@@

@@

m = rbind(get.roc.stat(p53roc),
      get.roc.stat(hgdpathroc),
      get.roc.stat(pathroc),
      get.roc.stat(roc))

@@

ggplot(melt(m), aes(model, value*100, group=variable, fill=variable)) + ylim(0,100) +
  geom_col(position="dodge") + 
  geom_text(aes(label=paste(round(value,2)*100,'%',sep='') ), position=position_dodge(width=1), vjust=-0.5) + 
  labs(x='', y='Percent', title='ROC') +
  scale_fill_manual(values=c('dodgerblue3','firebrick4', 'green4'), name='') + plot.theme + 
  theme(legend.position = 'bottom', axis.text.x=element_text(angle=45,hjust=1))

@@

@@

ggplot(m, aes(path, value, group=path, fill=variable)) + ylim(0,100) + facet_wrap(~status, ncol=1) +
  scale_fill_manual(values=c('firebrick4','green4', 'slateblue4'), name='', labels=c('Pred. progressed', 'Pred. Not progressed', 'p53 prediction'))  +
  geom_bar(stat='identity', position=position_stack()) +
  geom_text(aes(hjust=ifelse(value > 12, 2, 1),
                 label=paste(paste(round(value, 1),'%',sep=''),
                                        paste('n=', total, sep=''),sep='\n')), 
            position=position_stack(),color='white',  show.legend=F) +
  geom_bar(data=p53.melt, aes(path, value, group=path), fill='slateblue4',stat='identity',position = position_dodge()) +
  geom_text(data=p53.melt, aes(path,value,hjust=ifelse(value > 12, 2, 1),label=paste(paste(round(value, 1),'%',sep=''),paste('n=', total, sep=''),sep='\n')), position=position_dodge(),color='white',  show.legend=F) +
    labs(title='Samples predicted by pathology', x='Pathology', y='Percent', subtitle='Non-HGD exclude the final endoscopy, p53 excludes NA samples') +
  plot.theme + theme(text = element_text(size=16, face='bold'), legend.position = 'bottom') + coord_flip()

@@

## Flip order, 0-1 comes last
m = endos %>% group_by(brks, labels) %>% summarise( 'pred.r'= length(which(pred > 0))/length(brks), 'n'=length((Patient)) )

@@

ggplot(m, aes(brks, pred.r*100, group=brks)) + ylim(0,105) +
  geom_bar(stat='identity', fill='dodgerblue3', show.legend = F) +     
  scale_fill_manual(values=c('dodgerblue3', 'yellowgreen')) +
  geom_text(aes(y=pred.r*100+1.5, label=paste(round(pred.r*100), '%', sep=''))) +
  geom_text(aes(y=pred.r*100-2, label=paste('n=',n, sep='')), color='white') + 
  scale_x_discrete(labels=m$labels) + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = number of endoscopies') +
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1))

@@

## Progressor predictions per patient
```{r, echo=F, fig.height=12, fig.width=10}
plot.endo( subset(endos, Patient %in% subset(me, months <= -72)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(72/12, "years or more follow-up"))
```

```{r, echo=F, fig.height=12, fig.width=10}
# plot.endo( subset(endos, Patient %in% subset(me, months <= -48 & months > -72)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(48/12, "-",72/12,  "years or more follow-up"))
# plot.endo( subset(endos, Patient %in% subset(me, months <= -24 & months > -48)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(24/12, "-",48/12,  "years or more follow-up"))
plot.endo( subset(endos, Patient %in% subset(me, months > -48)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(1, "-",48/12,  "years or more follow-up"))
```

## Non-progressor predictions per patient
```{r, echo=F, fig.height=24, fig.width=16}
nonp.back = lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID],transform.pt.by.time)

@@

plot.endo(subset(nonp.endos, Patient %in% subset(nonp.endos, pred.ratio > 0)$Patient), bin=T) + labs(title='Non-progressors with progressive predictions')
```

@@

```{r spatial, echo=F}

# ids = subset(sum.patient.data, Status == 'P' & total.endos > 5 & M > mean(M, na.rm=T))$Hospital.Research.ID
# missing_p53 = do.call(rbind.data.frame, lapply(pg.samp, function(df) {
#   x = df[which(is.na(df$p53.Status)), c('Hospital.Research.ID', 'Path.ID', 'Status')]
#   if (nrow(x) > 0) x$Prioritize = ifelse(unique(df$Hospital.Research.ID) %in% ids, '*', '')
#   return(x)
# }))
# write.table(missing_p53, sep='\t', quote=F, row.names = F, col.names = T, file='missing_p53.txt')

#ids = subset(sum.patient.data, ((Status == 'P' & total.endos > 5) | (Status == 'NP' & total.endos > 9)) & M > mean(M, na.rm=T))$Hospital.Research.ID

transform.by.time<-function(pt) {
  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction')] %>% group_by(PID, Samplename) %>% mutate(
  'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12,
  'nl' = str_extract(Path.ID, "B\\w?[0-9]+"))

  pt$nl = (as.numeric(factor(pt$nl, levels=(unique(sort(pt$nl))), ordered = T))-1)*2
  #table(pt[,c('nl','months.before.final')])

@@

back = do.call(rbind.data.frame,lapply(pg.samp, transform.by.time))
```

```{r fig.width=8, fig.height=8}
probs = seq(0,1,1/4)
back$quants = with(back, cut(Prediction, breaks=quantile(Prediction, probs), include.lowest = T, labels=probs[-1]))

qt = table(back$quants, back$Status)

>>>>

```{r echo=F, warning=F, message=F, eval=T, fig.width=12, fig.height=8}

cn.mtn.plot<-function(df, label, col=c("orange4","grey90","steelblue3"), type='seg') {
  p = ggplot(df, aes(x=chr.length)) + facet_grid(Status~chr, scales='free_x', space='free_x', labeller=label)

  if (type == 'seg') {
    p = p + geom_segment(aes(x=start, xend=end, y=mean.value, yend=mean.value, color=cn), show.legend=T, size=1) + scale_color_manual(values=col) 
  }
  if (type == 'bar') {
    p = p + geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=cn), show.legend=T) + scale_fill_manual(values=col) 
  }

  p + labs(x='', y='Mean adjusted segmentation value', title='All samples') +
    theme(text=element_text(face='bold', size=14), axis.text.x=element_blank(), legend.position='bottom', 
          panel.grid = element_blank(), panel.background = element_blank(),strip.background =element_rect(fill="lightcyan2"),  
          panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines') )
}

chr.lengths = get.chr.lengths()
chr.lengths$chrom = sub('chr','',chr.lengths$chrom)
chr.lengths = subset(chr.lengths, chrom %in% c(1:22))
chr.lengths$chrom = factor(chr.lengths$chrom, levels=c(1:22), ordered = T)

arm.loc = arrange(rbind(
  (chr.lengths %>% group_by(chrom) %>% mutate('start'=1, 'end'=chr.cent-cent.gap))[c('chrom','start','end')],
  (chr.lengths %>% group_by(chrom) %>% mutate('start'=chr.cent+cent.gap, 'end'=chr.length))[c('chrom','start','end')]), chrom)

@@

globalCNdf = t(dysplasia.df[,grep('^\\d+:',colnames(dysplasia.df))])

locs = do.call(rbind.data.frame, lapply(rownames(globalCNdf), function(x) unlist(strsplit(x,':|-'))))
colnames(locs) = c('chr','start','end')
locs[] = lapply(locs[], function(x) as.numeric(as.character(x)))

globalCNdf = cbind.data.frame(locs, globalCNdf)
globalCNdf = melt(globalCNdf, id.vars=colnames(locs))

globalCNdf$seg.type = '5MB'
globalCNdf[with(globalCNdf, which( start %in% arm.loc$start & end %in% arm.loc$end)), 'seg.type'] = 'arms'
globalCNdf = merge(globalCNdf, chr.lengths[,c('chrom','chr.length')], by.x='chr', by.y='chrom')

globalCNdf = merge(globalCNdf, patient.info[,c('Hospital.Research.ID','Patient','Samplename','Pathology','Status')], by.x='variable', by.y='Samplename')
globalCNdf$Status = ifelse(globalCNdf$Status == 'P', 'Progressor', 'Non-Progressor')

cn = function(x) {
  str = 'neutral'
  if(x <= -1) {
    str = 'loss'
  } else if (x >= 1) {
    str = 'gain'
  }
  return(str)
}
globalCNdf$cn = factor(sapply(globalCNdf$value, cn), levels=c('loss','neutral','gain'), ordered = T)

@@

# Average cn value
globalCNdf = globalCNdf %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
globalCNdf$mean.value = globalCNdf$value # but not going to use it

globalCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value))
cnLegend = get.legend(globalCN.plot)
globalCN.plot = globalCN.plot + theme(legend.position = 'none')

@@

globalBECN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology == 'BE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')
#globalCNBEdf = subset(globalCNdf, Pathology == 'BE') %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )

@@

dfpts = subset(globalCNdf, Patient %in% c('57','76') & seg.type != 'arms')
#dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )

pt.status <- c('57'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
              '76'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
genomePtCN.plot = cn.mtn.plot(dfpts, labeller(Patient = pt.status))

dfpts = subset(globalCNdf, Patient %in% c('57','76') & Pathology == 'BE' & seg.type != 'arms')
#dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )

ptbe.status <- c('57'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
              '76'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))

genomePtBECN.plot = cn.mtn.plot(dfpts,  labeller(Patient = ptbe.status )) + labs(title='NDBE samples only')

genomePtCN.plot 
genomePtBECN.plot

cn.mtn.plot(dfpts,  labeller(Patient = ptbe.status ), type='bar') + labs(title='NDBE samples only')
```

## Individual samples
```{r}

a = sample(subset(patient.info, Status == 'NP' & Pathology %in% c('BE'))$Samplename, 1)
b = sample(subset(patient.info, Status == 'P' & Pathology %in% c('BE'))$Samplename, 1)

cn.mtn.plot(subset(globalCNdf, variable %in% c(a) & seg.type != 'arms'), labeller(label_both)) + labs(title="Individual NDBE samples") + theme(strip.text = element_text(size=10, face='bold'))
cn.mtn.plot(subset(globalCNdf, variable %in% c(b) & seg.type != 'arms'), labeller(label_both)) + labs(title="Individual NDBE samples") + theme(strip.text = element_text(size=10, face='bold'))

cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(label_both)) + labs(title="Individual NDBE samples")
cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(label_both), type='bar') + labs(title="Individual NDBE samples")

@@

```{r, echo=F, fig.height=5, fig.width=5}
mp = model.performance(performance.at.1se, coefs)
mp$plot + labs(title=paste('No HGD/IMC samples,', folds, 'folds,', splits, 'patient splits'))

@@

```{r, echo=F, fig.height=5, fig.width=5}
mp = model.performance(performance.at.1se, coefs)
mp$plot + labs(title=paste('No LGD/HGD/IMC samples,', folds, 'folds,', splits, 'patient splits'))

@@

ggplot(classification.rate, aes(model, mean, group=model, color=model)) + 
  geom_text( aes(label=round(mean, 3)), hjust=-0.5, size=5, color='gray39', show.legend = F) +
  geom_text( aes(y=mean+sme*2, label=paste(all.coef, "\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme),color='grey39', width=0.1, size=1) + 
  geom_point(size=4) +
  labs(x='', y='Classificaiton rate', title=paste('Classification rate per model at', select.alpha)) + 
  plot.theme + theme(axis.text.x=element_text(angle=45,hjust=1,face="bold",size=12), legend.position = 'none')

@@

custom.max<-function(cf, hazards, ymax) {
  if (cf > 0) {
    max = hazards*ymax
    max = ifelse(max > ymax, ymax, max)
  } else {
    max = hazards*ymax
    max = ifelse(max < ymax, ymax, max)
  }
  return(max)
}
chh = chh %>% rowwise() %>% mutate('max'=custom.max(coef, hazards, ymax))

globalCN.plot + geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)

globalBECN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)

ch$ymax = ifelse(ch$coef > 0, max(genomePtCN.plot$data$value), min(genomePtCN.plot$data$value))
chh = chh %>% rowwise() %>% mutate('max'=custom.max(coef, hazards, ymax))
genomePtCN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)

genomePtBECN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)

@@

```{r echo=F, fig.height=4, fig.width=8}
ggplot(chh, aes(x=label,y=coef,fill=gl)) + geom_bar(stat='identity') + labs(x='Effects', y='') +
  coord_flip() + scale_fill_manual(values=c('darkgreen','darkred')) +
  plot.theme + theme(axis.line.x=element_blank(), axis.ticks=element_blank(),axis.text=element_blank(), panel.grid=element_blank(),panel.grid.major=element_blank(), legend.position='none')
```

@@

```{r leaveoneout, echo=F, message=F, warning=F}
file = paste(cache.dir, 'loo.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop(paste("Missing model file", file))
}
loo.coefs = coefs

df = as.data.frame(performance.at.1se)
ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
  geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
  geom_label(data=subset(df, performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
  labs(y='performance at lambda.1se', x='', title=paste('Performance for nested leave one out, at alpha=',select.alpha,sep='')) +
  plot.theme

oddPt = names(pg.samp)[which.max(performance.at.1se)]
subset(sum.patient.data, Hospital.Research.ID == oddPt)

pg.samp[[oddPt]]$Prediction

# Not the pt with the most samples or endos
sum.patient.data[which.max(sum.patient.data$total.samples),]

@@

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=4, fig.width=4}
roc.plot <- function(roc,title="") {
  df = cbind.data.frame('Specificity'=rev(roc$specificities), 'Sensitivity'=rev(roc$sensitivities))

@@

```{r p53roc, echo=F, fig.height=4, fig.width=4}
p53samples = subset(patient.info, !is.na(p53.Status), select=c('Samplename','p53.Status','Status'))
p53roc = pROC::roc(as.numeric(p53samples$Status)-1, as.integer(p53samples$p53.Status)-1)

@@

```{r pathroc, echo=F, fig.height=4, fig.width=4 }
#sum(table(subset(patient.info, Status == 'P')$Pathology)[c('IMC','HGD')])/sum(table(subset(patient.info, Status == 'P')$Pathology))

@@

```{r echo=F, fig.height=6, fig.width=8}

@@

m = rbind.data.frame(get.roc.stat(p53roc),
      get.roc.stat(hgdpathroc),
      get.roc.stat(pathroc),
      get.roc.stat(roc))

@@

ggplot(melt(m,id.vars = 'model'), aes(model, value*100, group=variable, fill=variable)) + ylim(0,100) +
  geom_col(position="dodge") + 
  geom_text(aes(label=paste(round(value,2)*100,'%',sep='') ), position=position_dodge(width=1), vjust=-0.5) + 
  labs(x='', y='Percent', title='ROC') +
  scale_fill_manual(values=c('dodgerblue3','firebrick4', 'green4'), name='') + plot.theme + 
  theme(legend.position = 'bottom', axis.text.x=element_text(angle=45,hjust=1))

@@

## Jacknife coefficients

The idea here is to get some sort of CI for the coefficients, I'm not entirely sure this is correct though.

```{r jacknife, echo=F, warning=F, message=F}

coefs.per.pt = data.frame(matrix(ncol=length(pg.samp), nrow=nrow(ch), dimnames=list( ch$label, names(pg.samp))))
for (pt in names(pg.samp)) {
  lcf = loo.coefs[[pt]][,1]
  names(lcf) = rownames( loo.coefs[[pt]] )

  sharedcfs = intersect(names(lcf), ch$label)

  coefs.per.pt[sharedcfs, pt] = lcf[sharedcfs]
}
coefs.per.pt[is.na(coefs.per.pt)] = 0

library(bootstrap)

ch$jack.se = apply(coefs.per.pt, 1, function(x) {
  jk = jackknife(x, var)
  jk$jack.se
})

ch$adj = ifelse(ch$coef<0,-1,1)
ggplot(ch, aes(reorder(label, coef), coef)) + geom_point() + 
  geom_errorbar(aes(ymin=(abs(coef)-jack.se)*adj, ymax=(abs(coef)+jack.se)*adj)) + coord_flip() +
  plot.theme + labs(x='', title='Jackknifed coefficient CI??')

@@

@@

## Flip order, 0-1 comes last
m = endos %>% group_by(brks, labels) %>% summarise( 'pred.r'= length(which(pred > 0))/length(brks), 'n'=length((Patient)), 'n.samp'=sum(samples) )

@@

ggplot(m, aes(brks, pred.r*100, group=brks)) + ylim(0,105) +
  geom_bar(stat='identity', fill='dodgerblue3', show.legend = F) +     
  scale_fill_manual(values=c('dodgerblue3', 'yellowgreen')) +
  geom_text(aes(y=pred.r*100+1.5, label=paste(round(pred.r*100), '%', sep=''))) +
  geom_text(aes(y=pred.r*100-2, label=paste('n=',n,'\ns=',n.samp,sep='')), color='white') + 
  scale_x_discrete(labels=m$labels) + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1))

@@

## Progressor predictions per patient
```{r, echo=F, fig.height=8, fig.width=6}
#plot.endo( subset(endos, Patient %in% subset(me, months <= -72)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(72/12, "years or more follow-up"))
```

```{r, echo=F, fig.height=8, fig.width=6}
# plot.endo( subset(endos, Patient %in% subset(me, months <= -48 & months > -72)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(48/12, "-",72/12,  "years or more follow-up"))
# plot.endo( subset(endos, Patient %in% subset(me, months <= -24 & months > -48)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(24/12, "-",48/12,  "years or more follow-up"))
#plot.endo( subset(endos, Patient %in% subset(me, months > -48)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(1, "-",48/12,  "years or more follow-up"))
```

## Non-progressor predictions per patient
```{r, echo=F, fig.height=12, fig.width=8}
nonp.back = lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID],transform.pt.by.time)

@@

#plot.endo(subset(nonp.endos, Patient %in% subset(nonp.endos, pred.ratio > 0)$Patient), bin=T) + labs(title='Non-progressors with progressive predictions')
```

@@

```{r, spacial, fig.width=8, fig.height=6}
transform.by.time<-function(pt) {
  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction')] %>% group_by(PID, Samplename) %>% mutate(
  'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12
  #'nl' = Block
  #'nl' = str_extract(Path.ID, "B\\w?[0-9]+")
  )

  pt = merge(pt, subset(patient.info, Hospital.Research.ID == pt$Hospital.Research.ID[1], select=c('Samplename','Block')), by='Samplename')

  pt$nl = (pt$Block-1)*2
  #pt$nl = (as.numeric(factor(pt$nl, levels=(unique(sort(pt$nl))), ordered = T))-1)*2
  #table(pt[,c('nl','months.before.final')])

@@

back = do.call(rbind.data.frame,lapply(pg.samp, transform.by.time))
back = merge(back, patient.info[,c('Samplename','p53.Status')])

#cuts = seq(0,1,0.1)
cuts = quantile(back$Prediction, probs=seq(0,1,1/5))
back$quants = with(back, cut(Prediction, breaks=cuts))
qt = as.data.frame.matrix(table(back$quants, back$Status))
qt$quant = rownames(qt)

ft.fun<-function(NP,P) {
 f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
 #f$p.value
 cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
}

pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'conf'=ifelse(p.value < 0.05, '*', '') )
pred.confidence[c('r1','r2')] = NA
for (i in 1:(length(cuts)-1)) 
    pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)
